Application of Network Flow and Linear Programming to Scheduling of Clinicians

### 2019-07-05 paper/paper.intro.tex
Expand intro to describe our goals

Automated methods for tackling optimization problems related to scheduling have been studied extensively and applied in many industries, including transportation [??], manufacturing [??], [...]. Of special interest to our work are the approaches to tackle the nurse scheduling problem. In this problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences. A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??]. An extensive literature review of these and other methods is presented by [??]. We will briefly summarize the main ideas in some of these approaches. \\

[...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While this allows for finding precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians, it makes it very difficult to adapt a particular approach to other departments in the same hospital or other hospitals. Moreover, the large number of variables and constraints also leads to computational complexity issues, especially when using exact methods for finding the solution. \\

>>>>

Automated methods for tackling optimization problems related to scheduling have been studied extensively and applied in many industries, including transportation [??], manufacturing [??], [...]. Of special interest to our work are the approaches to tackle the nurse scheduling problem. In this problem, the goal is to find an optimal assignment of nurses to shifts that satisfies all of the hard constraints, such as hospital regulations, and as many soft constraints as possible, which may include nurse preferences. A wide variety of approaches, including exact and heuristic approaches, have been used to solve the nurse scheduling problem: integer linear programming [??], network flows [??], genetic algorithms [??], simulated annealing [??], and artificial intelligence [??]. An extensive literature review of these and other methods is presented by [??]. We will briefly summarize the main ideas of some of these approaches. \\

[...] Many of these approaches were designed to satisfy the requirements of a specific hospital department which causes a large number of variables and constraints to be incorporated into the problem formulation. While this allows for finding precise schedules that satisfy the needs of the department and the preferences of the nurses and clinicians, it makes it very difficult to adapt a particular approach to other departments in the same hospital or other hospitals. Moreover, the large number of variables and constraints also leads to computational complexity issues, especially when using exact methods for finding the solution. In this paper, we will tackle a version of the nurse scheduling problem arising from a case study of two on-call divisions of St. Michael's Hospital in Toronto, ON. Our goal is to present a simple formulation for the problem that has been developed and tested at the hospital after switching from a manual approach to scheduling, as well as analyze the performance of integer linear programming in solving difficult instances of the problem and compare the results with those of the manual approach. We hope that the formulation we present is adaptable, at least as a basic framework, for solving similar problems in other departments. \\
